---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```

# ptolemy: an R package for accessing global high-resolution geography

[![DOI](https://zenodo.org/badge/DOI/10.5281/zenodo.1470707.svg)](https://doi.org/10.5281/zenodo.1470707)

## About

This package was formerly known as `nPacMaps`. This R package relies on data downloaded
from the [Global Self-consistent, Hierarchical, High-resolution Geography (GSHHG) Database](https://www.ngdc.noaa.gov/mgg/shorelines/data/gshhg/latest/). More
details on the GSHHG can be found on this [website from the University of
Hawaii](http://www.soest.hawaii.edu/wessel/gshhg/). The current version 
supported in this package is 2.3.7 (July, 2017).

The main objective is to provide a simple interface for loading basemap
polygon land data for the North Pacific region that can be used in the
`sf` and `ggplot2` R package ecosystem. There are a variety of pre-described
regions that are loaded with convenience functions. However, most regions of
the world are supported and custom extractions are possible with `extract_gshhg()`.

This package is under development and functionality is subject to change
and improvement at anytime.

Please cite this package as:

> London, Josh M. ptolemy: an R package for accessing global high-resolution geography. 2018. Version 1.0.0. https://github.com/jmlondon/ptolemy. 10.5281/zenodo.1470707

and, also cite the GSHHG dataset:

> Wessel, P., and W. H. F. Smith. A Global Self-consistent, Hierarchical, High-resolution Shoreline Database. J. Geophys. Res., 101, 8741-8743, 1996.

## Installation

The ptolemy package is not available on CRAN and must be installed via the
`devtools::install_github()` function.

```{r, install-code, eval=FALSE}
install.packages("devtools")
devtools::install_github('jmlondon/ptolemy')
```

After successfully installing the package from GitHub, you will need to download
and install the GSHGG data. This is handled via the `ptolemy::install_gshhg()`
function. If you are not prompted, you may need to run `ptolemy::install_gsggh()`
before using the package.

```{r, install-gshgg, eval=FALSE}
library(ptolemy)
install_gshhg()
```

## Data for Specifying the Geographic Region

The `data` parameter should be an `sf` object representing the region of interest.
This can be data points, polygons, lines and should be projected. If the data
are provided in longlat, an *epsg* parameter code is required.

## Resolution

The GSHHG dataset has five different resolutions available:

1. **f**ull resolution: Original (full) data resolution.
2. **h**igh resolution: About 80 % reduction in size and quality.
3. **i**ntermediate resolution: Another ~80 % reduction.
4. **l**ow resolution: Another ~80 % reduction.
5. **c**rude resolution: Another ~80 % reduction.

The **i**ntermediate reolustion has been set as the default option and should
suffice for most applications. The default resolution can be overided via the
`resolution` parameter. Users should consider bumping up to **h**igh or
**f**ull when zooming into smaller scale regions. This will increase the 
extraction and drawing time. If you require the **f**ull resolution frequently, 
creating a custom region via `extract_gshhg()` should be considered.

## ESPG / Projection

All of the returned maps are provide with projected coordinates based on the 
`epsg` parameter provided. For the pre-built regions, the default projections
are sensible and all users need to do is insure all other data is transformed
to the same projection. A custom projection can be provided.

## Buffer

The returned map region will be slightly larger than the area of interest
represented by the provided data. The default is 5000 meters.

## Simplified for Improved Performance

The `rmapshaper::ms_simplify` function can applied to the returned objects to 
improve performance. The function is set to return 20% of the original points 
and to preserve small shapes (i.e. small islands and other features). This 
improves performance for plotting and should suffice for most users. However, 
the original integrity of the GSHHG data will be changed. Uses for analytical 
purposes (e.g. calculating distance to shoreline, least-cost path creation 
around land) should not use this option without consideration of impacts.

## geom_sf for ggplot2

By default all of the basemap objects are returned as an `sf` object. ggplot2
now offers native support for plotting `sf` objects via the `geom_sf`. The 
examples below demonstrate basic use of `geom_sf`.

## Acknowledgements

This package relies heavily on the works of other researchers and developers.
The GSHHG dataset is developed and maintained by Paul Wessel (SOEST, University 
of Hawai'i, Honolulu, HI) and Walter H. F. Smith (NOAA Geosciences Lab, 
National Ocean Service, Silver Spring, MD). The package also relies on code
developed as part of the `PBSmapping` package (Jon T. Schnute, Nicholas Boers 
and Rowan Haigh (2018). PBSmapping: Mapping Fisheries Data and Spatial Analysis 
Tools. R package version 2.70.5. https://CRAN.R-project.org/package=PBSmapping.)

## Examples

### North Pacific Basemap

```{r, npac-example}
library(ggplot2)
library(ptolemy)
library(sf)

npac_base <- ptolemy::npac()

npac_plot <- ggplot() + 
  geom_sf(data = npac_base,
               fill = "grey60", size = 0.2) +
  ggtitle('North Pacific Basemap (epsg:3832)')
npac_plot
```

### California Current

```{r, calcur-example}
library(ggplot2)
library(ptolemy)
library(sf)

calcur_base <- ptolemy::calcur()

calcur_plot <- ggplot() + 
  geom_sf(data = calcur_base,
               fill = "grey60", size = 0.2) +
  ggtitle('California Current Basemap (epsg:3310)')
calcur_plot
```

### Bering Sea Basemap

```{r, bering-example}
library(ggplot2)
library(ptolemy)
library(sf)

bering_base <- ptolemy::bering()

bering_plot <- ggplot() + 
  geom_sf(data = bering_base,
               fill = "grey60", size = 0.2) +
  ggtitle('Bering Sea Basemap (epsg:3571)')
bering_plot
```

### US (Alaska) Arctic Basemap

```{r, us-arctic-example}
library(ggplot2)
library(ptolemy)
library(sf)

us_arctic_base <- ptolemy::us_arctic()

us_arctic_plot <- ggplot() + 
  geom_sf(data = us_arctic_base,
               fill = "grey60", size = 0.2) +
  ggtitle('US Arctic Basemap (epsg:3572)')
us_arctic_plot
```

### Alaska Basemap

```{r, ak-example}
library(ggplot2)
library(ptolemy)
library(sf)

ak_base <- ptolemy::alaska()

ak_plot <- ggplot() + 
  geom_sf(data = ak_base,
               fill = "grey60", size = 0.2) +
  ggtitle('Alaska Basemap (epsg:3338)')
ak_plot
```

We can also zoom in on a particular region

```{r, ak-example-zoom}
library(ggplot2)
library(ptolemy)
library(crawl)
library(dplyr)
library(sf)

data("harborSeal")

harborSeal <- harborSeal %>% 
  filter(!is.na(latitude)) %>% 
  as.data.frame() %>% 
  sf::st_as_sf(coords = c("longitude","latitude")) %>% 
  sf::st_set_crs(4326) %>% 
  sf::st_transform(3338)


map_base <- ptolemy::extract_gshhg(data = harborSeal)

ak_plot <- ggplot() + 
  geom_sf(data = map_base,
               fill = "grey60", size = 0.2) +
  geom_sf(data = harborSeal,
             alpha = 0.1, color = 'blue') 
ak_plot
```

*************************************************

##### Disclaimer

<sub>This repository is a scientific product and is not official communication of the Alaska Fisheries Science Center, the National Oceanic and Atmospheric Administration, or the United States Department of Commerce. All AFSC Marine Mammal Laboratory (AFSC-MML) GitHub project code is provided on an ‘as is’ basis and the user assumes responsibility for its use. AFSC-MML has relinquished control of the information and no longer has responsibility to protect the integrity, confidentiality, or availability of the information. Any claims against the Department of Commerce or Department of Commerce bureaus stemming from the use of this GitHub project will be governed by all applicable Federal law. Any reference to specific commercial products, processes, or services by service mark, trademark, manufacturer, or otherwise, does not constitute or imply their endorsement, recommendation or favoring by the Department of Commerce. The Department of Commerce seal and logo, or the seal and logo of a DOC bureau, shall not be used in any manner to imply endorsement of any commercial product or activity by DOC or the United States Government.</sub>

