---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```

# nPacMaps: an R package for North Pacific basemap data

## About

executive summary of the package and functions

## Installation

The nPacMaps package is not available on CRAN and must be installed via the
`devtools::install_github()` function.

```{r, install-code, eval=FALSE}
install.packages("devtools")
devtools::install_github('jmlondon/npacmaps',ref='develop')
```

After successfully installing the package from GitHub, you will need to download
and install the GSHGG data. This is handled via the `nPacMaps::install_gshhg()`
function.

```{r, install-gshgg, eval=FALSE}
library(nPacMaps)
install_gshhg()
```

## Examples

### North Pacific Basemap

```{r, npac-example}
library(ggplot2)
library(nPacMaps)

npac_base <- nPacMaps::npac()

npac_plot <- ggplot() + 
  geom_polygon(data = npac_base,
               aes(x = long,y = lat,group = group,id = id),
               fill = "grey60") +
  coord_fixed() +
  xlab("easting (km)") + ylab("northing (km)") +
  scale_x_continuous(labels = nPacMaps::to_km()) +
  scale_y_continuous(labels = nPacMaps::to_km())
npac_plot
```

### Alaska Basemap

```{r, ak-example}
library(ggplot2)
library(nPacMaps)

ak_base <- nPacMaps::alaska()

ak_plot <- ggplot() + 
  geom_polygon(data = ak_base,
               aes(x = long,y = lat,group = group,id = id),
               fill = "grey60") +
  coord_fixed() +
  xlab("easting (km)") + ylab("northing (km)") +
  scale_x_continuous(labels = nPacMaps::to_km()) +
  scale_y_continuous(labels = nPacMaps::to_km())
ak_plot
```

We can also zoom in on a particular region

```{r, ak-example-zoom}
library(ggplot2)
library(nPacMaps)
library(crawl)
library(dplyr)

data("harborSeal")

harborSeal <- harborSeal %>% filter(!is.na(latitude)) %>% 
  as.data.frame()

sp::coordinates(harborSeal) <- ~longitude + latitude
sp::proj4string(harborSeal) <- CRS("+init=epsg:4326")

harborSeal <- sp::spTransform(harborSeal, CRS("+init=epsg:3338"))
harborSeal <- as.data.frame(harborSeal)

map_limits <- nPacMaps::ggExpansion(harborSeal,x = "longitude",y = "latitude",
                                    x_fac = 0.25, y_fac = 0.25)

ak_base <- nPacMaps::alaska()

ak_plot <- ggplot() + 
  geom_polygon(data = ak_base,
               aes(x = long,y = lat,group = group,id = id),
               fill = "grey60") +
  geom_point(data = harborSeal,aes(x = longitude, y = latitude), 
             alpha = 0.1, color = 'blue') +
  coord_fixed(xlim = map_limits$xlim, ylim = map_limits$ylim) +
  xlab("easting (km)") + ylab("northing (km)") +
  scale_x_continuous(labels = nPacMaps::to_km()) +
  scale_y_continuous(labels = nPacMaps::to_km())
ak_plot
```

