---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```

# nPacMaps: an R package for North Pacific basemap data

## About

Creating maps for the North Pacific can be difficult, frustrating, and confusing.
The main source of this frustration is the 180 longitude line and the limited
examples that exist for guidance. This R package relies on data from the [Global 
Self-consistent, Hierarchical, High-resolution Geography (GSHHG) Database](https://www.soest.hawaii.edu/pwessel/gshhg/).

The main objective is to provide a simple interface for quickly loading basemap
polygon land data for the North Pacific Region that can be used in either the
ggplot2 or sp:spplot graphical ecosystems.

This package is under active development and functionality is subject to change
and improvement at anytime.

## Installation

The nPacMaps package is not available on CRAN and must be installed via the
`devtools::install_github()` function.

```{r, install-code, eval=FALSE}
install.packages("devtools")
devtools::install_github('jmlondon/npacmaps')
```

A cutting-edge 'develop' branch is also available. Proceed with caution.

```{r, install-code-develop, eval=FALSE}
install.packages("devtools")
devtools::install_github('jmlondon/npacmaps',ref='develop')
```

After successfully installing the package from GitHub, you will need to download
and install the GSHGG data. This is handled via the `nPacMaps::install_gshhg()`
function.

```{r, install-gshgg, eval=FALSE}
library(nPacMaps)
install_gshhg()
```

## Resolution

The GSHHG dataset has five different resolutions available:

1. **f**ull resolution: Original (full) data resolution.
2. **h**igh resolution: About 80 % reduction in size and quality.
3. **i**ntermediate resolution: Another ~80 % reduction.
4. **l**ow resolution: Another ~80 % reduction.
5. **c**rude resolution: Another ~80 % reduction.

The **i**ntermediate reolustion has been set as the default option and should
suffice for most applications. The default resolution can be overided via the
`resolution` parameter. Users should consider bumping up to **h**igh or
**f**ull when zooming into smaller scale regions. This will increase the 
extraction and drawing time. If you require the **f**ull resolution frequently, 
creating a custom region via `extract_gshhg()` should be considered.

## Projections

All of the returned maps are provide with projected coordinates in a default
projection chosen for each region. In most cases, the default projections
are sensible and all users need to do is insure all other data is transformed
to the same projection. A custom projection can be provided by specifying the 
corresponding epsg integer value (as a character) to the `epsg` parameter.

## Fortify for ggplot2

By default all of the basemap objects are returned as a `data.frame` that has
been _fortified_ via the `broom::tidy()` function. If you would like, instead,
for the object to be returned as a _SpatialPolygonsDataFrame_ for use with `plot()`
or `sp::spplot()` then the `fortify` parameter should be specified as `FALSE`.

## Additional Utility Functions

The package also includes two additional utility functions: `to_km()` and 
`ggExpansion()`. The `to_km()` function provides simple conversion of the
easting and northing labels from meters to kilometers. The `ggExpansion()`
function allows the user to easily specify an expansion fraction for the xlim and
ylim components of the ggplot.

## Examples

### North Pacific Basemap

```{r, npac-example}
library(ggplot2)
library(nPacMaps)

npac_base <- nPacMaps::npac()

npac_plot <- ggplot() + 
  geom_polygon(data = npac_base,
               aes(x = long,y = lat,group = group,id = id),
               fill = "grey60") +
  coord_fixed() +
  xlab("easting (km)") + ylab("northing (km)") +
  scale_x_continuous(labels = nPacMaps::to_km()) +
  scale_y_continuous(labels = nPacMaps::to_km()) +
  ggtitle('North Pacific Basemap (epsg:3832)')
npac_plot
```

### California Current

```{r, calcur-example}
library(ggplot2)
library(nPacMaps)

calcur_base <- nPacMaps::calcur()

calcur_plot <- ggplot() + 
  geom_polygon(data = calcur_base,
               aes(x = long,y = lat,group = group,id = id),
               fill = "grey60") +
  coord_fixed() +
  xlab("easting (km)") + ylab("northing (km)") +
  scale_x_continuous(labels = nPacMaps::to_km()) +
  scale_y_continuous(labels = nPacMaps::to_km()) +
  ggtitle('California Current Basemap (epsg:3310)')
calcur_plot
```

### Bering Sea Basemap

```{r, bering-example}
library(ggplot2)
library(nPacMaps)

bering_base <- nPacMaps::bering()

bering_plot <- ggplot() + 
  geom_polygon(data = bering_base,
               aes(x = long,y = lat,group = group,id = id),
               fill = "grey60") +
  coord_fixed() +
  xlab("easting (km)") + ylab("northing (km)") +
  scale_x_continuous(labels = nPacMaps::to_km()) +
  scale_y_continuous(labels = nPacMaps::to_km()) +
  ggtitle('Bering Sea Basemap (epsg:3571)')
bering_plot
```

### US (Alaska) Arctic Basemap

```{r, us-arctic-example}
library(ggplot2)
library(nPacMaps)

us_arctic_base <- nPacMaps::us_arctic()

us_arctic_plot <- ggplot() + 
  geom_polygon(data = us_arctic_base,
               aes(x = long,y = lat,group = group,id = id),
               fill = "grey60") +
  coord_fixed() +
  xlab("easting (km)") + ylab("northing (km)") +
  scale_x_continuous(labels = nPacMaps::to_km()) +
  scale_y_continuous(labels = nPacMaps::to_km()) +
  ggtitle('US Arctic Basemap (epsg:3572)')
us_arctic_plot
```

### Alaska Basemap

```{r, ak-example}
library(ggplot2)
library(nPacMaps)

ak_base <- nPacMaps::alaska()

ak_plot <- ggplot() + 
  geom_polygon(data = ak_base,
               aes(x = long,y = lat,group = group,id = id),
               fill = "grey60") +
  coord_fixed() +
  xlab("easting (km)") + ylab("northing (km)") +
  scale_x_continuous(labels = nPacMaps::to_km()) +
  scale_y_continuous(labels = nPacMaps::to_km()) +
  ggtitle('Alaska Basemap (epsg:3338)')
ak_plot
```

We can also zoom in on a particular region

```{r, ak-example-zoom}
library(ggplot2)
library(nPacMaps)
library(crawl)
library(dplyr)

data("harborSeal")

harborSeal <- harborSeal %>% filter(!is.na(latitude)) %>% 
  as.data.frame()

sp::coordinates(harborSeal) <- ~longitude + latitude
sp::proj4string(harborSeal) <- CRS("+init=epsg:4326")

harborSeal <- sp::spTransform(harborSeal, CRS("+init=epsg:3338"))
harborSeal <- as.data.frame(harborSeal)

map_limits <- nPacMaps::ggExpansion(harborSeal,x = "longitude",y = "latitude",
                                    x_fac = 0.25, y_fac = 0.25)

ak_base <- nPacMaps::alaska(resolution = "f")

ak_plot <- ggplot() + 
  geom_polygon(data = ak_base,
               aes(x = long,y = lat,group = group,id = id),
               fill = "grey60") +
  geom_point(data = harborSeal,aes(x = longitude, y = latitude), 
             alpha = 0.1, color = 'blue') +
  coord_fixed(xlim = map_limits$xlim, ylim = map_limits$ylim) +
  xlab("easting (km)") + ylab("northing (km)") +
  scale_x_continuous(labels = nPacMaps::to_km()) +
  scale_y_continuous(labels = nPacMaps::to_km())
ak_plot
```

*************************************************

#####Disclaimer

<sub>This repository is a scientific product and is not official communication of the Alaska Fisheries Science Center, the National Oceanic and Atmospheric Administration, or the United States Department of Commerce. All AFSC Marine Mammal Laboratory (AFSC-MML) GitHub project code is provided on an ‘as is’ basis and the user assumes responsibility for its use. AFSC-MML has relinquished control of the information and no longer has responsibility to protect the integrity, confidentiality, or availability of the information. Any claims against the Department of Commerce or Department of Commerce bureaus stemming from the use of this GitHub project will be governed by all applicable Federal law. Any reference to specific commercial products, processes, or services by service mark, trademark, manufacturer, or otherwise, does not constitute or imply their endorsement, recommendation or favoring by the Department of Commerce. The Department of Commerce seal and logo, or the seal and logo of a DOC bureau, shall not be used in any manner to imply endorsement of any commercial product or activity by DOC or the United States Government.</sub>

